---
title: "Take-home Exercise 4"
description: |
  This is my Take-home Exercise to reveal the daily routine of two residents in Ohio, USA.
author:
  - name: Joyce Tseng
    url: https://www.linkedin.com/in/joyce-tseng-a7115a1aa/
    affiliation: School of Computing and Information Systems (SMU)
    affiliation_url: https://scis.smu.edu.sg/master-it-business
date: "`r Sys.Date()`"
preview:
output:
  distill::distill_article:
    code_folding: Show code
    self_contained: false
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE, fig.retina = 3)
```

# 1. Overview

In this take-home exercise, we reveal the daily routine of the most happiness and the less happiness residents in Ohio to see the differences on their everyday behavior. The data is provided and can be downloaded from [VAST Challenge 2022](https://vast-challenge.github.io/2022/). The package of ***VisiElse*** is the main method to plot the graph.  

# 2. Data Preparation
### 2.1 Installing Packages

```{r}
packages = c('tidyverse', 'plotly', 'dplyr', 'data.table', 'lubridate', 'zoo',
             'ViSiElse', 'scales', 'viridis', 'ggthemes')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

### 2.2 Importing Data

```{r eval=FALSE}
log1 <- read_csv("ParticipantStatusLogs1.csv")
log2 <- read_csv("ParticipantStatusLogs2.csv")
participants <- read_csv("Participants.csv")

# merge two logs together
df <- rbind(log1, log2)
```

### 2.3 Data Wrangling

***PICKING TWO RESIDENTS***

Here we would like to pick two residents who have the highest (joviality = 0.999234) and the lowest (joviality = 0.000204) value of joviality in Ohio to observe their daily action.

```{r eval=FALSE, include=FALSE}
saveRDS(df, file = "df.rds")
saveRDS(participants, file = "participants.rds")
```

```{r include=FALSE}
participants <- readRDS(file = "participants.rds")
```

```{r}
# Get max and min joviality of participants_id
# slice_max and slice_min is function from the tidyverse
max <- participants %>% 
  slice_min(joviality)
min <- participants %>% 
  slice_max(joviality)

max
min
```

***FILTERING PICKED RESIDENTS***

```{r eval=FALSE}
# filter two participants
df <- df %>% 
  filter(participantId == "758" | participantId == "131") 
```

***DERIVING TIME FIELDS***

To derive time relative data like date, day and hour, we use the functions from `lubridate` and `zoo` to extract them.

```{r eval=FALSE}
# extract date, weekday, hour, minute from timestamp
df <- df %>%
  mutate(date = as.Date(timestamp)) %>%
  mutate(day = weekdays(date)) %>%
  mutate(hour = hour(timestamp)) %>%
  mutate(min = minute(timestamp)) %>%
  mutate(minutes = hour*60 + min)
```

***DROPPING UNNEEDED DATA***

The data in activity logs records participants' daily behaviour from 2022-03-01 to 2023-05-31, but we just pick the first week (2022-03-01 to 2022-03-07) of data to plot in our exercise.

```{r eval=FALSE}
df <- df %>%
  filter(date < "2022-03-08")
```

```{r eval=FALSE, include=FALSE}
saveRDS(df, file = "df.rds")
```

```{r include=FALSE}
df <- readRDS(file = "df.rds")
```

***CREATING WEEKLY DATASETS FOR SLEEP, EATEN AND HUNGER***

```{r}
# Weekly sleep data
sleep <- df %>% 
  select(participantId, date, day, sleepStatus, hour, min, minutes) %>%
  filter(sleepStatus == "Sleeping")

# Weekly eaten data
eaten <- df %>% 
  select(participantId, date, day, hungerStatus, hour, min, minutes) %>%
  filter(hungerStatus == "JustAte")

# Weekly hunger data
hunger <- df %>% 
  select(participantId, date, day, hungerStatus, hour, min, minutes) %>%
  filter(hungerStatus == "Hungry")
```

```{r eval=FALSE, include=FALSE}
# Weekly sleep status
week_sleep <- df  %>% 
  select(participantId, date, day, sleepStatus, hour, min, minutes) %>% 
  group_by(participantId, sleepStatus) %>%
  dplyr::summarise(value = min(minutes))

week_sleep <- pivot_wider(week_sleep, names_from = sleepStatus, values_from = value)

# Weekly hunger status
week_hunger <- df %>% 
  select(participantId, date, day, hungerStatus, hour, min, minutes) %>% 
  group_by(participantId, hungerStatus) %>%
  dplyr::summarise(value = min(minutes))
  
week_hunger <- pivot_wider(week_hunger, names_from = hungerStatus, values_from = value)

# Weekly mode
week_mode <- df %>% 
  select(participantId, date, day, currentMode, hour, min, minutes)%>% 
  group_by(participantId, currentMode) %>%
  dplyr::summarise(value = min(minutes))

week_mode <- pivot_wider(week_mode, names_from = currentMode, values_from = value)

# Merger three behavior
week <- merge(x = week_sleep, y = week_hunger, by = "participantId", all = TRUE)
week <- merge(x = week, y = week_mode, by = "participantId", all = TRUE)
week
```

***CREATING DAILY DATASET***

```{r}
# create only Monday data
Monday <- df %>% 
  select(participantId, date, day, currentMode, hungerStatus,
         sleepStatus, hour, min, minutes) %>%
  filter(date == "2022-03-07")

# create only Tuesday data
Tuesday <- df %>% 
  select(participantId, date, day, currentMode, hungerStatus,
         sleepStatus, hour, min, minutes) %>%
  filter(date == "2022-03-01")

# create only Wednesday data
Wednesday <- df %>% 
  select(participantId, date, day, currentMode, hungerStatus,
         sleepStatus, hour, min, minutes) %>%
  filter(date == "2022-03-02")

# create only Thursday data
Thursday <- df %>% 
  select(participantId, date, day, currentMode, hungerStatus,
         sleepStatus, hour, min, minutes) %>%
  filter(date == "2022-03-03")

# create only Friday data
Friday <- df %>% 
  select(participantId, date, day, currentMode, hungerStatus,
         sleepStatus, hour, min, minutes) %>%
  filter(date == "2022-03-04")

# create only Saturday data
Saturday <- df %>% 
  select(participantId, date, day, currentMode, hungerStatus,
         sleepStatus, hour, min, minutes) %>%
  filter(date == "2022-03-05")

# create only Sunday data
Sunday <- df %>% 
  select(participantId, date, day, currentMode, hungerStatus,
         sleepStatus, hour, min, minutes) %>%
  filter(date == "2022-03-06")
```

# 3. Data Visualization
In our dataset, although the daily behavior includes punctual and long action, we would just focus on the start time of all doings. Thus, all actions are regarded as short action.

### 3.1 Daily Pattern on Monday 

***PIVOTTING AND MERGERING THE DATA***

```{r}
# Monday sleep status
Mon_sleep <- Monday %>% 
  group_by(participantId, sleepStatus) %>%
  dplyr::summarise(value = min(minutes))

Mon_sleep <- pivot_wider(Mon_sleep, names_from = sleepStatus, values_from = value)

# Monday hunger status
Mon_hunger <- Monday %>% 
  group_by(participantId, hungerStatus) %>%
  dplyr::summarise(value = min(minutes))
  
Mon_hunger <- pivot_wider(Mon_hunger, names_from = hungerStatus, values_from = value)

# Monday mode
Mon_mode <- Monday %>% 
  group_by(participantId, currentMode) %>%
  dplyr::summarise(value = min(minutes))

Mon_mode <- pivot_wider(Mon_mode, names_from = currentMode, values_from = value)

# Merger three behavior
Mon <- merge(x = Mon_sleep, y = Mon_hunger, by = "participantId", all = TRUE)
Mon <- merge(x = Mon, y = Mon_mode, by = "participantId", all = TRUE)
```

***ARRANGING THE COLUMN ORDER***

```{r}
# rearrange the columns
Mon_order <- c("participantId", "Sleeping", "Awake", "BecomingHungry", "JustAte",
           "Transport", "AtWork", "Hungry", "AtRestaurant", "AtRecreation", 
           "AtHome")
Mon <- Mon[,Mon_order] 
```

```{r include=FALSE}
visi1 <- visielse(Mon, pixel = 30)
book <- ConvertFromViSibook(visi1@book)
```

***VISUALIZATION***

```{r}
# Monday
group <- c("Id: 113", "Id: 758")
visi2 <- visielse(Mon,
                  book = book,
                  group = group,
                  method = "cut",
                  tests = F,
                  pixel = 30,
                  doplot = F)
plot(visi2, vp0w = 0.7, unit.tps = "min", scal.unit.tps = 30)
```

### 3.2 Daily Pattern on Tuesday 

***PIVOTTING AND MERGERING THE DATA***

```{r}
# Tuesday sleep status
Tue_sleep <- Tuesday %>% 
  group_by(participantId, sleepStatus) %>%
  dplyr::summarise(value = min(minutes))

Tue_sleep <- pivot_wider(Tue_sleep, names_from = sleepStatus, values_from = value)

# Tuesday hunger status
Tue_hunger <- Tuesday %>% 
  group_by(participantId, hungerStatus) %>%
  dplyr::summarise(value = min(minutes))
  
Tue_hunger <- pivot_wider(Tue_hunger, names_from = hungerStatus, values_from = value)

# Tuesday mode
Tue_mode <- Tuesday %>% 
  group_by(participantId, currentMode) %>%
  dplyr::summarise(value = min(minutes))

Tue_mode <- pivot_wider(Tue_mode, names_from = currentMode, values_from = value)

# Merger three behavior
Tue <- merge(x = Tue_sleep, y = Tue_hunger, by = "participantId", all = TRUE)
Tue <- merge(x = Tue, y = Tue_mode, by = "participantId", all = TRUE)
```

***ARRANGING THE COLUMN ORDER***

```{r}
# rearrange the columns
Tue_order <- c("participantId", "Sleeping", "Awake", "BecomingHungry", "JustAte",
           "BecameFull", "Transport", "AtWork", "Hungry", "AtRestaurant", "AtRecreation", 
           "AtHome", "PrepareToSleep")
Tue <- Tue[,Tue_order] 
```

***VISUALIZATION***

```{r}
# Tuesday
group <- c("Id: 113", "Id: 758")
visi3 <- visielse(Tue,
                  book = book,
                  group = group,
                  method = "cut",
                  tests = F,
                  pixel = 30,
                  doplot = F)
plot(visi3, vp0w = 0.7, unit.tps = "min", scal.unit.tps = 30)
```

### 3.3 Daily Pattern on Wednesday 

***PIVOTTING AND MERGERING THE DATA***

```{r}
# Wednesday sleep status
Wed_sleep <- Wednesday %>% 
  group_by(participantId, sleepStatus) %>%
  dplyr::summarise(value = min(minutes))

Wed_sleep <- pivot_wider(Wed_sleep, names_from = sleepStatus, values_from = value)

# Wednesday hunger status
Wed_hunger <- Wednesday %>% 
  group_by(participantId, hungerStatus) %>%
  dplyr::summarise(value = min(minutes))
  
Wed_hunger <- pivot_wider(Wed_hunger, names_from = hungerStatus, values_from = value)

# Wednesday mode
Wed_mode <- Wednesday %>% 
  group_by(participantId, currentMode) %>%
  dplyr::summarise(value = min(minutes))

Wed_mode <- pivot_wider(Wed_mode, names_from = currentMode, values_from = value)

# Merger three behavior
Wed <- merge(x = Wed_sleep, y = Wed_hunger, by = "participantId", all = TRUE)
Wed <- merge(x = Wed, y = Wed_mode, by = "participantId", all = TRUE)
```

***ARRANGING THE COLUMN ORDER***

```{r}
# rearrange the columns
Wed_order <- c("participantId", "Sleeping", "Awake", "BecomingHungry", "Hungry", "JustAte",
           "BecameFull", "Transport", "AtWork", "Hungry", "AtRestaurant", "AtRecreation", 
           "AtHome", "PrepareToSleep")
Wed <- Wed[,Wed_order] 
```

***VISUALIZATION***

```{r}
# Wednesday
group <- c("Id: 113", "Id: 758")
visi4 <- visielse(Wed,
                  book = book,
                  group = group,
                  method = "cut",
                  tests = F,
                  pixel = 30,
                  doplot = F)
plot(visi4, vp0w = 0.7, unit.tps = "min", scal.unit.tps = 30)
```

### 3.4 Daily Pattern on Thursday 

***PIVOTTING AND MERGERING THE DATA***

```{r}
# Thursday sleep status
Thu_sleep <- Thursday %>% 
  group_by(participantId, sleepStatus) %>%
  dplyr::summarise(value = min(minutes))

Thu_sleep <- pivot_wider(Thu_sleep, names_from = sleepStatus, values_from = value)

# Thursday hunger status
Thu_hunger <- Thursday %>% 
  group_by(participantId, hungerStatus) %>%
  dplyr::summarise(value = min(minutes))
  
Thu_hunger <- pivot_wider(Thu_hunger, names_from = hungerStatus, values_from = value)

# Thursday mode
Thu_mode <- Thursday %>% 
  group_by(participantId, currentMode) %>%
  dplyr::summarise(value = min(minutes))

Thu_mode <- pivot_wider(Thu_mode, names_from = currentMode, values_from = value)

# Merger three behavior
Thu <- merge(x = Thu_sleep, y = Thu_hunger, by = "participantId", all = TRUE)
Thu <- merge(x = Thu, y = Thu_mode, by = "participantId", all = TRUE)
```

***ARRANGING THE COLUMN ORDER***

```{r}
# rearrange the columns
Thu_order <- c("participantId", "Sleeping", "Awake", "BecomingHungry", "Hungry", "JustAte",
           "BecameFull", "Transport", "AtWork", "Hungry", "AtRestaurant", "AtRecreation", 
           "AtHome", "PrepareToSleep")
Thu <- Thu[,Thu_order] 
```

***VISUALIZATION***

```{r}
# Thursday
group <- c("Id: 113", "Id: 758")
visi5 <- visielse(Thu,
                  book = book,
                  group = group,
                  method = "cut",
                  tests = F,
                  pixel = 30,
                  doplot = F)
plot(visi5, vp0w = 0.7, unit.tps = "min", scal.unit.tps = 30)
```

### 3.5 Daily Pattern on Friday 

***PIVOTTING AND MERGERING THE DATA***

```{r}
# Friday sleep status
Fri_sleep <- Friday %>% 
  group_by(participantId, sleepStatus) %>%
  dplyr::summarise(value = min(minutes))

Fri_sleep <- pivot_wider(Fri_sleep, names_from = sleepStatus, values_from = value)

# Friday hunger status
Fri_hunger <- Friday %>% 
  group_by(participantId, hungerStatus) %>%
  dplyr::summarise(value = min(minutes))
  
Fri_hunger <- pivot_wider(Fri_hunger, names_from = hungerStatus, values_from = value)

# Friday mode
Fri_mode <- Friday %>% 
  group_by(participantId, currentMode) %>%
  dplyr::summarise(value = min(minutes))

Fri_mode <- pivot_wider(Fri_mode, names_from = currentMode, values_from = value)

# Merger three behavior
Fri <- merge(x = Fri_sleep, y = Fri_hunger, by = "participantId", all = TRUE)
Fri <- merge(x = Fri, y = Fri_mode, by = "participantId", all = TRUE)
```

***ARRANGING THE COLUMN ORDER***

```{r}
# rearrange the columns
Fri_order <- c("participantId", "Sleeping", "Awake", "BecomingHungry", "Hungry", "JustAte",
           "BecameFull", "Transport", "AtWork", "Starving", "AtRestaurant", "AtRecreation", 
           "AtHome", "PrepareToSleep")
Fri <- Fri[,Fri_order] 
```

***VISUALIZATION***

```{r}
# Friday
group <- c("Id: 113", "Id: 758")
visi6 <- visielse(Fri,
                  book = book,
                  group = group,
                  method = "cut",
                  tests = F,
                  pixel = 30,
                  doplot = F)
plot(visi6, vp0w = 0.7, unit.tps = "min", scal.unit.tps = 30)
```

### 3.6 Daily Pattern on Saturday 

***PIVOTTING AND MERGERING THE DATA***

```{r}
# Saturday sleep status
Sat_sleep <- Saturday %>% 
  group_by(participantId, sleepStatus) %>%
  dplyr::summarise(value = min(minutes))

Sat_sleep <- pivot_wider(Sat_sleep, names_from = sleepStatus, values_from = value)

# Saturday hunger status
Sat_hunger <- Saturday %>% 
  group_by(participantId, hungerStatus) %>%
  dplyr::summarise(value = min(minutes))
  
Sat_hunger <- pivot_wider(Sat_hunger, names_from = hungerStatus, values_from = value)

# Saturday mode
Sat_mode <- Saturday %>% 
  group_by(participantId, currentMode) %>%
  dplyr::summarise(value = min(minutes))

Sat_mode <- pivot_wider(Sat_mode, names_from = currentMode, values_from = value)

# Merger three behavior
Sat <- merge(x = Sat_sleep, y = Sat_hunger, by = "participantId", all = TRUE)
Sat <- merge(x = Sat, y = Sat_mode, by = "participantId", all = TRUE)
```

***ARRANGING THE COLUMN ORDER***

```{r}
# rearrange the columns
Sat_order <- c("participantId", "Sleeping", "Awake", "BecomingHungry", "Hungry", "JustAte",
           "BecameFull", "Transport", "AtWork", "Starving", "AtRestaurant", "AtRecreation", 
           "AtHome", "PrepareToSleep")
Sat <- Sat[,Sat_order] 
```

***VISUALIZATION***

```{r}
# Saturday
group <- c("Id: 113", "Id: 758")
visi7 <- visielse(Sat,
                  book = book,
                  group = group,
                  method = "cut",
                  tests = F,
                  pixel = 30,
                  doplot = F)
plot(visi7, vp0w = 0.7, unit.tps = "min", scal.unit.tps = 30)
```

```{r eval=FALSE, include=FALSE}
# Sunday sleep status
Sun_sleep <- Sunday %>% 
  group_by(participantId, sleepStatus) %>%
  dplyr::summarise(value = min(minutes))

Sun_sleep <- pivot_wider(Sun_sleep, names_from = sleepStatus, values_from = value)

# Sunrday hunger status
Sun_hunger <- Sunday %>% 
  group_by(participantId, hungerStatus) %>%
  dplyr::summarise(value = min(minutes))
  
Sun_hunger <- pivot_wider(Sun_hunger, names_from = hungerStatus, values_from = value)

# Saturday mode
Sun_mode <- Sunday %>% 
  group_by(participantId, currentMode) %>%
  dplyr::summarise(value = min(minutes))

Sun_mode <- pivot_wider(Sun_mode, names_from = currentMode, values_from = value)

# Merger three behavior
Sun <- merge(x = Sun_sleep, y = Sun_hunger, by = "participantId", all = TRUE)
Sun <- merge(x = Sun, y = Sun_mode, by = "participantId", all = TRUE)
Sun
```

```{r eval=FALSE, include=FALSE}
# rearrange the columns
Sun_order <- c("participantId", "Sleeping", "Awake", "BecomingHungry", "Hungry", "JustAte",
           "BecameFull", "Transport", "AtWork", "Starving", "AtRestaurant", 
           "AtHome", "PrepareToSleep")
Sun <- Sun[,Sun_order] 
```

```{r eval=FALSE, include=FALSE}
# Sunday
group <- c("Id: 113", "Id: 758")
visi8 <- visielse(Sun,
                  book = book,
                  group = group,
                  method = "cut",
                  tests = F,
                  pixel = 30,
                  doplot = F)
plot(visi8, vp0w = 0.7, unit.tps = "min", scal.unit.tps = 30)
```

### 3.7 Insight on Everday Pattern

The order of daily action has been arranged ascendingly a person's normal one day schedule. We can observe that participant id = 113 and id = 758 would go to sleep before 24:00. The awake time on each day for this two people is quite different and even at midnight. It is because that they may wake up to go to the toilet or feel thirsty to drink water in the middle of night. Therefore, the awake time sometimes is not in the morning.

In addition, it is also can observable that their everyday working time is also fluctuating, meaning that they may arrive and start to work a bit earlier or later. After leaving their workplace, it is more often for participant id = 758 to go to the restaurant to have a dinner and have recreational activities. This may relates to his low joviality, since he probably needs some ways to release his negative emotion or pressure.

### 3.8 Weekly Pattern on Sleeping

```{r}
grouped <- sleep %>% 
  count(day, hour) %>% 
  ungroup() %>%
  na.omit()
  
ggplot(grouped, 
       aes(hour, 
           day, 
           fill = n)) + 
geom_tile(color = "white", 
          size = 0.1) + 
theme_tufte(base_family = "Helvetica") + 
coord_equal() +
scale_fill_gradient(name = "Sleeping Time",
                    low = "sky blue", 
                    high = "dark blue") +
labs(x = NULL, 
     y = NULL, 
     title = "Weekly Sleeping Time") +
theme(axis.ticks = element_blank(),
      plot.title = element_text(hjust = 0.5),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6) )
```

A regular sleeping time can be seen from the graph. There is no differences wake-up time and sleep time on weekday and weekend. Normally, they would go to sleep at around 11pm to 12pm and wake up at 9am.

### 3.9 Weekly Pattern on Eaten

```{r}
grouped <- eaten %>% 
  count(day, hour) %>% 
  ungroup() %>%
  na.omit()
  
ggplot(grouped, 
       aes(hour, 
           day, 
           fill = n)) + 
geom_tile(color = "white", 
          size = 0.1) + 
theme_tufte(base_family = "Helvetica") + 
coord_equal() +
scale_fill_gradient(name = "Eaten Time",
                    low = "sky blue", 
                    high = "dark blue") +
labs(x = NULL, 
     y = NULL, 
     title = "Weekly Eaten Time") +
theme(axis.ticks = element_blank(),
      plot.title = element_text(hjust = 0.5),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6) )
```

The eaten time for people each day is quite different, but we can observe that normally they would have 3 meals in one day and prefer having their lunch a bit late. 

### 3.10 Weekly Pattern on Hunger

```{r}
grouped <- hunger %>% 
  count(day, hour) %>% 
  ungroup() %>%
  na.omit()
  
ggplot(grouped, 
       aes(hour, 
           day, 
           fill = n)) + 
geom_tile(color = "white", 
          size = 0.1) + 
theme_tufte(base_family = "Helvetica") + 
coord_equal() +
scale_fill_gradient(name = "Hunger Time",
                    low = "sky blue", 
                    high = "dark blue") +
labs(x = NULL, 
     y = NULL, 
     title = "Weekly Hunger Time") +
theme(axis.ticks = element_blank(),
      plot.title = element_text(hjust = 0.5),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6) )
```

The graph of pattern on eaten and hunger is more or less complementary. It can be seen that when there is no record on eaten time, then the color would present in huger graph. For example, during midnight and afternoon.







